---
title: "Federal Student Aid Per Student Analysis"
output: html_document
---
## Creating Loops for Processing Data for Loans 2010-12

```{r temp, results="hide", message=FALSE, warning=FALSE}

# Loading Libraries

library(readxl)
library(tidyr)
library(stringr)
library(dplyr)
library(DT)

# Setting Up Directory and Gathering a List

list_files <- list.files(path = "CSV.Files")

list_files_loans_6 <- list_files <- list.files(path = "CSV.Files/Loans.6")

# Looping for Bringing in Excel Files

# Loan.6 Looping

## Loop Length

loan_length_6 <- length(list_files_loans_6)
for (i in 1:loan_length_6) {

## i is what cycle the loop is on
  
loan_filename_6 <- list_files_loans_6[i] 

## Choosing Excel Files to Loop and Creating Path Names

path_name_loans_6 <-  paste0("CSV.Files/Loans.6/", loan_filename_6)
print(path_name_loans_6)

temp_dataframe_6 <- read_excel(path_name_loans_6, sheet = 2, col_names = TRUE, col_types = NULL, na = "", skip = 5) 

## Changing Column Names

colnames(temp_dataframe_6) <- c("OPE.ID", "School", "State", "Zip.Code", "School.Type", "Recipients.1", "Loans.No.1", "Loans.Sum.1", "Disbursements.No.1", "Disbursements.Sum.1", "Recipients.2", "Loans.No.2", "Loans.Sum.2", "Disbursements.No.2", "Disbursements.Sum.2","Recipients.3", "Loans.No.3", "Loans.Sum.3", "Disbursements.No.3", "Disbursements.Sum.3", "Recipients.4", "Loans.No.4", "Loans.Sum.4", "Disbursements.No.4", "Disbursements.Sum.4", "Recipients.5", "Loans.No.5", "Loans.Sum.5", "Disbursements.No.5", "Disbursements.Sum.5",
"Recipients.6", "Loans.No.6", "Loans.Sum.6", "Disbursements.No.6", "Disbursements.Sum.6")

temp_dataframe_6 <- temp_dataframe_6[c("School", "State", "School.Type", "Disbursements.Sum.1", "Disbursements.Sum.2", "Disbursements.Sum.3", "Disbursements.Sum.4", "Disbursements.Sum.5", "Disbursements.Sum.6")]
                                   
temp_dataframe_6$Total.Disbursements <- temp_dataframe_6$Disbursements.Sum.1 + temp_dataframe_6$Disbursements.Sum.2 + temp_dataframe_6$Disbursements.Sum.3 + temp_dataframe_6$Disbursements.Sum.4 + temp_dataframe_6$Disbursements.Sum.5 + temp_dataframe_6$Disbursements.Sum.6

temp_dataframe_6 <- temp_dataframe_6[c("School", "State", "School.Type", "Total.Disbursements")]

temp_dataframe_6$Year <- loan_filename_6

temp_dataframe_6$Year <- gsub("DL_Dashboard_AY", "", temp_dataframe_6$Year)
temp_dataframe_6$Year <- gsub("_Q4.xls", "", temp_dataframe_6$Year)

if (i == 1) {
mega_dataframe_6 <- temp_dataframe_6  
} else (
  mega_dataframe_6 <- rbind(mega_dataframe_6, temp_dataframe_6)
)

}

```

### Table for Loans 2010-12
```{r table, warning=FALSE}

datatable(mega_dataframe_6)

```




## Creating Loops for Processing Data for Loans 2012-16
```{r temp2, results="hide", message=FALSE, warning=FALSE}
# Gathering a List

list_files <- list.files(path = "CSV.Files")

list_files_loans_5 <- list_files <- list.files(path = "CSV.Files/Loans.5")

# Loan.5 Looping

## Loop Length

loan_length_5 <- length(list_files_loans_5)
for (i in 1:loan_length_5) {
  
## i is what cycle the loop is on
  
loan_filename_5 <- list_files_loans_5[i] 
  
## Choosing Excel Files to Loop and Creating Path Names
  
path_name_loans_5 <- paste0("CSV.Files/Loans.5/", loan_filename_5)
print(path_name_loans_5)
  
temp_dataframe_5 <- read_excel(path_name_loans_5, sheet = 2, col_names = TRUE, col_types = NULL, na = "", skip = 5) 
  
## Changing Column Names
  
colnames(temp_dataframe_5) <- c("OPE.ID", "School", "State", "Zip.Code", "School.Type", "Recipients.1", "Loans.No.1", "Loans.Sum.1", "Disbursements.No.1", "Disbursements.Sum.1", "Recipients.2", "Loans.No.2", "Loans.Sum.2", "Disbursements.No.2", "Disbursements.Sum.2", "Recipients.3", "Loans.No.3", "Loans.Sum.3", "Disbursements.No.3", "Disbursements.Sum.3", "Recipients.4", "Loans.No.4", "Loans.Sum.4", "Disbursements.No.4", "Disbursements.Sum.4", "Recipients.5", "Loans.No.5", "Loans.Sum.5", "Disbursements.No.5", "Disbursements.Sum.5")
                                
temp_dataframe_5 <- temp_dataframe_5[c("School", "State", "School.Type", "Disbursements.Sum.1", "Disbursements.Sum.2", "Disbursements.Sum.3", "Disbursements.Sum.4", "Disbursements.Sum.5")]

temp_dataframe_5$Total.Disbursements <- temp_dataframe_5$Disbursements.Sum.1 + temp_dataframe_5$Disbursements.Sum.2 + temp_dataframe_5$Disbursements.Sum.3 + temp_dataframe_5$Disbursements.Sum.4 + temp_dataframe_5$Disbursements.Sum.5

temp_dataframe_5 <- temp_dataframe_5[c("School", "State", "School.Type", "Total.Disbursements")]
  
temp_dataframe_5$Year <- loan_filename_5
  
temp_dataframe_5$Year <- gsub("DL_Dashboard_AY", "", temp_dataframe_5$Year)
temp_dataframe_5$Year <- gsub("_Q4.xls", "", temp_dataframe_5$Year)
  
if (i == 1) {
mega_dataframe_5 <- temp_dataframe_5  
} else (
  mega_dataframe_5 <- rbind(mega_dataframe_5, temp_dataframe_5)
)

}

```

### Table for Loans 2012-16
```{r table2, warning=FALSE}

datatable(mega_dataframe_5)

```



## Creating Loops for Processing Data for Grants 2010-11
```{r temp4, results="hide", message=FALSE, warning=FALSE}
# Grant.5 Looping

## File Names

list_files_grants_5 <- list_files <- list.files(path = "CSV.files/Grants.5")

## Loop Length

grant_length_5 <- length(list_files_grants_5)
  
## i is what cycle the loop is on
  
grant_filename_5 <- list_files_grants_5 
  
## Choosing Excel Files to Loop and Creating Path Names
  
path_name_grants_5 <- paste0("CSV.Files/Grants.5/", grant_filename_5)
print(path_name_grants_5)
  
temp_dataframe.5 <- read_excel(path_name_grants_5, sheet = 3, col_names = TRUE, col_types = NULL, na = "", skip = 5) 
## Changing Column Names
  
colnames(temp_dataframe.5) <- make.names(colnames(temp_dataframe.5))
  
colnames(temp_dataframe.5) <- c("OPE.ID", "School", "State", "Zip.Code", "School.Type", "YTD.Recipients.1", "YTD.Disbursements.1", "YTD.Recipients.2", "YTD.Disbursements.2", "YTD.Recipients.3", "YTD.Disbursements.3", "YTD.Recipients.4", "YTD.Disbursements.4", "YTD.Recipients.5", "YTD.Disbursements.5") 
  
temp_dataframe.5 <- temp_dataframe.5[c("School", "State", "School.Type", "YTD.Disbursements.1", "YTD.Disbursements.2", "YTD.Disbursements.3", "YTD.Disbursements.4", "YTD.Disbursements.5")]
  
temp_dataframe.5$Total.Disbursements <- temp_dataframe.5$YTD.Disbursements.1 + temp_dataframe.5$YTD.Disbursements.2 + temp_dataframe.5$YTD.Disbursements.3 + temp_dataframe.5$YTD.Disbursements.4 + temp_dataframe.5$YTD.Disbursements.5
  
temp_dataframe.5 <- temp_dataframe.5[c("School", "State", "School.Type", "Total.Disbursements")]

temp_dataframe.5$Year <- grant_filename_5
  
temp_dataframe.5$Year <- gsub("Q4", "", temp_dataframe.5$Year)
temp_dataframe.5$Year <- gsub("AY.xls", "", temp_dataframe.5$Year)
  
```

### Table for Grants 2010-11
```{r table4, warning=FALSE}

datatable(temp_dataframe.5)

```




## Creating Loops for Processing Data for Grants 2011-16
```{r temp3, results="hide", message=FALSE, warning=FALSE}

# Grant.3 Looping

## File Names

list_files_grants_3 <- list_files <- list.files(path = "CSV.files/Grants.3")

## Loop Length

grant_length_3 <- length(list_files_grants_3)
for (i in 1:grant_length_3) {
  
## i is what cycle the loop is on
  
grant_filename_3 <- list_files_grants_3[i] 
  
## Choosing Excel Files to Loop and Creating Path Names
  
path_name_grants_3 <-  paste0("CSV.Files/Grants.3/", grant_filename_3)
print(path_name_grants_3)
  
temp_dataframe_3 <- read_excel(path_name_grants_3, sheet = 3, col_names = TRUE, col_types = NULL, na = "", skip = 5) 
## Changing Column Names

colnames(temp_dataframe_3) <- make.names(colnames(temp_dataframe_3))
  
colnames(temp_dataframe_3) <- c("OPE.ID", "School", "State", "Zip.Code", "School.Type", "YTD.Recipients.1", "YTD.Disbursements.1", "YTD.Recipients.2", "YTD.Disbursements.2", "YTD.Recipients.3", "YTD.Disbursements.3")

temp_dataframe_3 <- temp_dataframe_3[c("School", "State", "School.Type", "YTD.Disbursements.1", "YTD.Disbursements.2", "YTD.Disbursements.3")]
  
temp_dataframe_3$Total.Disbursements <- temp_dataframe_3$YTD.Disbursements.1 + temp_dataframe_3$YTD.Disbursements.2 + temp_dataframe_3$YTD.Disbursements.3
  
temp_dataframe_3 <- temp_dataframe_3[c("School", "State", "School.Type", "Total.Disbursements")]
  
temp_dataframe_3$Year <- grant_filename_3
  
temp_dataframe_3$Year <- gsub("Q4", "", temp_dataframe_3$Year)
temp_dataframe_3$Year <- gsub("AY.xls", "", temp_dataframe_3$Year)
  
if (i == 1) {
mega_dataframe_3 <- temp_dataframe_3
} else (
  mega_dataframe_3 <- rbind(mega_dataframe_3, temp_dataframe_3)
)
  
}

```

### Table for Grants 2011-16
```{r table3, warning=FALSE}

datatable(mega_dataframe_3)

```


## Creating Loops for Processing Data for Campus-Based Programs 2010-15
```{r temp5, results="hide", message=FALSE, warning=FALSE}

# Gathering a List

list_files <- list.files(path = "CSV.Files")

list_files_campus <- list_files <- list.files(path = "CSV.Files/Campus")

campus_length <- length(list_files_campus)
for (i in 1:campus_length) {
  
## i is what cycle the loop is on
  
campus_filename <- list_files_campus[i] 

## Choosing Excel Files to Loop and Creating Path Names
  
path_name_campus <-  paste0("CSV.Files/Campus/", campus_filename)
print(path_name_campus)

temp_dataframe_campus <- read_excel(path_name_campus, sheet = 1, col_names = TRUE, col_types = NULL, na = "", skip = 3) 
  
  
## Changing Column Names
  
colnames(temp_dataframe_campus) <- c("OPE.ID", "School", "State", "Zip.Code", "School.Type", "Recipients.1", "Federal.Award.1", "Disbursements.1", "Recipients.2", "Federal.Award.2", "Disbursements.2", "Recipients.3", "Federal.Award.3", "Disbursements.3")
  
temp_dataframe_campus <- temp_dataframe_campus[c("School", "State", "School.Type", "Disbursements.1", "Disbursements.2", "Disbursements.3")]
  
temp_dataframe_campus$Total.Disbursements <- temp_dataframe_campus$Disbursements.1 + temp_dataframe_campus$Disbursements.2 + temp_dataframe_campus$Disbursements.3
  
temp_dataframe_campus <- temp_dataframe_campus[c("School", "State", "School.Type", "Total.Disbursements")]
  
temp_dataframe_campus$Year <- campus_filename
  
temp_dataframe_campus$Year <- gsub("CBData.xls", "", temp_dataframe_campus$Year)
 
if (i == 1) {
mega_dataframe_campus <- temp_dataframe_campus  
} else (
  mega_dataframe_campus <- rbind(mega_dataframe_campus, temp_dataframe_campus)
)
  
}

```


### Table for Campus-Based Programs 2010-15

```{r table5, warning=FALSE}

datatable(mega_dataframe_campus)

```




## Totaling Federal Student Aid

```
I exported the dataframes above and renamed them when I imported them.
```

```{r temp6, results="hide", message=FALSE, warning=FALSE}

# Importing CSV Files

grants.10.11 <- read.csv(file = "grants.10-11.csv")
grants.11.16 <- read.csv(file = "grants.11-16.csv")
loans.10.12 <- read.csv(file = "loans.10-12.csv")
loans.12.16 <- read.csv(file = "loans.12-16.csv")
campus.10.15 <- read.csv(file = "campus.10-15.csv")

# Joining

loans.10.16 <- full_join(loans.10.12, loans.12.16)
grants.10.16 <- full_join(grants.10.11, grants.11.16)

# Grouping

loan.state.summary <- loans.10.16 %>%
  group_by(State, School.Type, Year) %>%
  summarise(Sum.Type.loan=sum(Total.Disbursements))

grant.state.summary <- grants.10.16 %>%
  group_by(State, School.Type, Year) %>%
  summarise(Sum.Type.grant=sum(Total.Disbursements))

campus.state.summary <- campus.10.15 %>%
  group_by(State, School.Type, Year) %>%
  summarise(Sum.Type.campus=sum(Total.Disbursements))

loan.state.summary <- subset(loan.state.summary, !is.na(loan.state.summary$State))
grant.state.summary <- subset(grant.state.summary, !is.na(grant.state.summary$State))
campus.state.summary <- subset(campus.state.summary, !is.na(campus.state.summary$State))

# Fixing Cases

loan.state.summary$School.Type <- str_to_title(loan.state.summary$School.Type)

grant.state.summary$School.Type <- str_to_title(grant.state.summary$School.Type)

campus.state.summary$School.Type <- str_to_title(campus.state.summary$School.Type)

loan.state.summary$School.Type <- gsub("Private-Nonprofit", "Private", loan.state.summary$School.Type)

grant.state.summary$School.Type <- gsub("Private-Nonprofit", "Private", grant.state.summary$School.Type)

campus.state.summary$School.Type <- gsub("Private/Non-Profit", "Private", campus.state.summary$School.Type)

loan.state.summary$School.Type <- gsub("Foreign*", "", loan.state.summary$School.Type)

grant.state.summary$School.Type <- gsub("Foreign*", "", grant.state.summary$School.Type)

campus.state.summary$School.Type <- gsub("Foreign*", "", campus.state.summary$School.Type)

loan.state.summary$Year <- gsub("2010_2011", "1011", loan.state.summary$Year)

loan.state.summary$Year <- gsub("2011_2012", "1112", loan.state.summary$Year)

loan.state.summary$Year <- gsub("2012_2013", "1213", loan.state.summary$Year)

loan.state.summary$Year <- gsub("2013_2014", "1314", loan.state.summary$Year)

loan.state.summary$Year <- gsub("2014_2015", "1415", loan.state.summary$Year)

loan.state.summary$Year <- gsub("2015_2016", "1516", loan.state.summary$Year)

campus.state.summary$Year <- gsub("2010-11", "1011", campus.state.summary$Year)

campus.state.summary$Year <- gsub("2011-12", "1112", campus.state.summary$Year)

campus.state.summary$Year <- gsub("2012-13", "1213", campus.state.summary$Year)

campus.state.summary$Year <- gsub("2013-14", "1314", campus.state.summary$Year)

campus.state.summary$Year <- gsub("2014-15", "1415", campus.state.summary$Year)

campus.state.summary$Year <- str_trim(campus.state.summary$Year)

grant.state.summary$Year <- as.character(grant.state.summary$Year)

campus.state.summary$Year <- as.character(campus.state.summary$Year)

# Joining

loans_grants_join <- full_join(loan.state.summary, grant.state.summary)

# Subsetting

loans_grants_join <- subset(loans_grants_join, !is.na(loans_grants_join$State))

loans_grants_join <- subset(loans_grants_join, State!="FC")
loans_grants_join <- subset(loans_grants_join, State!="PR")
loans_grants_join <- subset(loans_grants_join, State!="GU")
loans_grants_join <- subset(loans_grants_join, State!="AS")
loans_grants_join <- subset(loans_grants_join, State!="FM")
loans_grants_join <- subset(loans_grants_join, State!="MH")
loans_grants_join <- subset(loans_grants_join, State!="MP")
loans_grants_join <- subset(loans_grants_join, State!="VI")
loans_grants_join <- subset(loans_grants_join, State!="PW")

loans_grants_join$Total.Disbursement <- loans_grants_join$Sum.Type.loan + loans_grants_join$Sum.Type.grant

campus.state.summary <- subset(campus.state.summary, !is.na(campus.state.summary$State))

campus.state.summary <- subset(campus.state.summary, State!="FC")
campus.state.summary <- subset(campus.state.summary, State!="PR")
campus.state.summary <- subset(campus.state.summary, State!="GU")
campus.state.summary <- subset(campus.state.summary, State!="AS")
campus.state.summary <- subset(campus.state.summary, State!="FM")
campus.state.summary <- subset(campus.state.summary, State!="MH")
campus.state.summary <- subset(campus.state.summary, State!="MP")
campus.state.summary <- subset(campus.state.summary, State!="VI")
campus.state.summary <- subset(campus.state.summary, State!="PW")

# Exporting

loans_grants_join <- write.csv(loans_grants_join, "loans.grants.join.csv")
campus.state.summary <- write.csv(campus.state.summary, "campus.csv")

```

```
I had to export the files and manually join them in Excel and then import the finalized version because the 
campus-based programs did not have data for the year 2015-16 and some proprietary schools were missing.
```

```{r temp8, results="hide", message=FALSE, warning=FALSE}

# Importing

loans.grants.campus.join <- read.csv(file = "loans.grants.campus.join.csv")

# Fixing Cases

loans.grants.campus.join$State <- str_trim(loans.grants.campus.join$State)
loans.grants.campus.join$Total.Disbursement <- as.numeric(loans.grants.campus.join$Total.Disbursement)

# Final Dataframe

loans.grants.campus.total <- loans.grants.campus.join %>%
  dplyr::select(State, School.Type, Year, Total.Disbursement) %>%
  group_by(State, Year) %>%
  mutate(Year.Total=sum(Total.Disbursement), Percent=round(Total.Disbursement/Year.Total, 4)) %>%
  select(State, School.Type, Year, Percent) %>%
spread(Year, Percent)


```


### Table for Total Federal Student Aid

```{r table6, warning=FALSE}

datatable(loans.grants.campus.total)

```


## Creating Looping and Totaling Enrollment

```{r temp7, results="hide", message=FALSE, warning=FALSE}

# Gathering a List

list.files(path = "CSV.Files/Enrollment")
list_files <- list.files(path = "CSV.Files/Enrollment")

# Looping for Bringing in Excel Files

# Enrollment Looping

## File Names

list_files_enroll <- list_files <- list.files(path = "CSV.Files/Enrollment")

## Loop Length

enroll_length <- length(list_files_enroll)
for (i in 1:enroll_length) {
  
## i is what cycle the loop is on
  
enroll_filename <- list_files_enroll[i] 

## Choosing Excel Files to Loop and Creating Path Names
  
path_name_enroll <-  paste0("CSV.Files/Enrollment/", enroll_filename)
print(path_name_enroll)
  
  temp_enroll <- read_excel(path_name_enroll, sheet = 1, col_names = TRUE, col_types = NULL, na = "") 
  
  ## Changing Column Names
  
  colnames(temp_enroll) <- make.names(colnames(temp_enroll))
  
  temp_enroll$Year <- enroll_filename
  
  temp_enroll$Year <- gsub("2010-11.enroll.CT.xls", "2010-11", temp_enroll$Year)
  temp_enroll$Year <- gsub("2011-12.enroll.CT.xls", "2011-12", temp_enroll$Year)
  temp_enroll$Year <- gsub("2012-13.enroll.CT.xls", "2012-13", temp_enroll$Year)
  temp_enroll$Year <- gsub("2013-14.enroll.CT.xls", "2013-14", temp_enroll$Year)
  
  if (i == 1) {
    mega_enroll <- temp_enroll 
  } else (
    mega_enroll <- rbind(mega_enroll, temp_enroll)
  )

}

## Exporting

write.csv(mega_enroll, "enroll.10.14.final.csv")

# Importing

enroll.10.14 <- read.csv(file = "enroll.10.14.csv")

college.name <- read_excel(path = "college.names.final.xlsx", sheet = 1, col_names = TRUE, col_types = NULL, na = "")


# Changing Column Names

enroll.10.14 <- enroll.10.14[c("School", "Enrollment", "Year")]
college.name <- college.name[c("School", "Type")]
college.name <- subset(college.name, !is.na(college.name$School))

# Fixing Cases

college.name$School <- str_to_title(college.name$School)
enroll.10.14$School <- str_to_title(enroll.10.14$School)
college.name$School <- str_trim(college.name$School)
enroll.10.14$School <- str_trim(enroll.10.14$School)
enroll.10.14$School <- as.character(enroll.10.14$School)

# Joining

enroll.join <- full_join(enroll.10.14, college.name)

# Cleaning

enroll.join <- subset(enroll.join, !is.na(enroll.join$Type))
enroll.join <- subset(enroll.join, Enrollment!="NA")
enroll.join$Enrollment <- as.numeric(enroll.join$Enrollment)

# Final Dataframe

enroll.final <- enroll.join %>%
  dplyr::select(School, Enrollment, Type, Year) %>%
  group_by(Year, Type) %>%
  summarise(Enroll.Total=sum(Enrollment, na.rm=T)) %>%
  mutate(Year.Total=sum(Enroll.Total), Percent=round(Enroll.Total/Year.Total, 4)) %>%
  select(Type, Year, Percent) %>%
  spread(Year, Percent)

```


### Table for Connecticut Enrollment

```{r table7, warning=FALSE}

datatable(enroll.final)

```
